Abstraction and Invariance for Algebraically Indexed Types

### 2012-10-24 paper/geometry-twocolumn.tex
summary figure in geometry
Text about figure 2

Linear maps of the form $\left(
  \begin{smallmatrix}s & 0 \\ 0 & s\end{smallmatrix}\right)$, as
generated by $\iota_1$, commute with all other invertible linear
maps. We thus require
$(\iota_1 s)B = B(\iota_1 s)$ to hold.
The scaling maps $\left(
  \begin{smallmatrix}s & 0 \\ 0 & s\end{smallmatrix}\right)$
are precisely the elements of $\GL{2}$ that commute with all others;
these form the \emph{centre} of $\GL{2}$. If we keep track of
scalings, then we can assign the following more precise types to
scalar multiplication and dot product:
\begin{displaymath}
\begin{array}{@{}l@{}l}
  (*) : \forall s \mathord: \SynGL{1}, B \mathord:
  \SynGL{2}.\ & \tyPrim{real}{s} \to \tyPrim{vec}{B,0} \to
  \tyPrim{vec}{\iota_1(s)B,0}\\
%\end{displaymath}
%and dot product:
%\begin{displaymath}
%  \begin{array}{l@{}l}
    (\cdot) : \forall s \mathord: \SynGL{1}, O \mathord: \SynOrth{2}.\ & \tyPrim{vec}{\iota_1(s)\iota_O(O),0} \to\\
    &\tyPrim{vec}{\iota_1(s)\iota_O(O),0} \to \tyPrim{real}{s^2}
  \end{array}
\end{displaymath}

@@

>>>>

Linear maps of the form $\left(
  \begin{smallmatrix}s & 0 \\ 0 & s\end{smallmatrix}\right)$, as
generated by $\iota_1$, commute with all other invertible linear
maps. We thus require
$(\iota_1 s)B = B(\iota_1 s)$ to hold.
The scaling maps $\left(
  \begin{smallmatrix}s & 0 \\ 0 & s\end{smallmatrix}\right)$
are precisely the elements of $\GL{2}$ that commute with all others;
these form the \emph{centre} of $\GL{2}$. If we keep track of
scalings, then we can assign the more precise types to
scalar multiplication and dot product. These
 are shown in Figure~\ref{fig:vec-ops}, which summarises the most general types 
of all the vector operations that we have described.

@@

\begin{figure}[t]
  \centering
  \begin{align*}
    0 & : \forall B \mathord: \SynGL{2}.\ \tyPrim{vec}{B,0} \\
    (+) & :
    \begin{array}[t]{@{}l}
      \forall B \mathord: \SynGL{2}, t_1,t_2 \mathord: \SynTransl{2}. \\
      \tyPrim{vec}{B,t_1} \to \tyPrim{vec}{B,t_2} \to
      \tyPrim{vec}{B,t_1 + t_2}
    \end{array}
    \\
    \mathrm{negate} & : \forall B \mathord: \SynGL{2}, t \mathord: \SynTransl{2}.\ \tyPrim{vec}{B,t} \to \tyPrim{vec}{B,-t} \\
  (*) & : 
  \begin{array}[t]{@{}l}
    \forall s \mathord: \SynGL{1}, B \mathord:\SynGL{2}.\ \\
    \tyPrim{real}{s} \to \tyPrim{vec}{B,0} \to \tyPrim{vec}{\iota_1(s)B,0}
  \end{array}\\
    \mathrm{affComb} &:
    \begin{array}[t]{@{}l}\forall t \mathord:
    \SynTransl{2}.\ \tyPrim{vec}{t} \to \tyPrimNm{real} \to
    \tyPrim{vec}{t} \to \tyPrim{vec}{t} \\ 
    \sem{\mathrm{affComb}}\ \vec{v_1}\ r\ \vec{v_2} = (1-r)\vec{v_1} +
    r\vec{v_2} 
    \end{array}
    (\cdot) & :
\begin{array}[t]{@{}l}
    \forall s \mathord: \SynGL{1}, O \mathord: \SynOrth{2}.\ \\
    \tyPrim{vec}{\iota_1(s)\iota_O(O),0} \to\tyPrim{vec}{\iota_1(s)\iota_O(O),0} \to \tyPrim{real}{s^2}
\end{array}\\
  (\times) & : \forall B \mathord: \SynGL{2}.\ \tyPrim{vec}{B,0} \to
  \tyPrim{vec}{B,0} \to \tyPrim{real}{\det B} 
\end{align*}
  \caption{Operations on vectors}
  \label{fig:vec-ops}
\end{figure}
